00-Data-Processing-for-Food-Banks-Accessibility

### 2021-04-20 00-Data-Processing/00-Data-Processing-for-Food-Banks-Accessibility.Rmd

We can get an API key to use `cancensus` from [CensusMapper](https://censusmapper.ca/). Once you have an API key, it can be stored locally this way:
```{r eval=FALSE}
set_api_key("your-key", install = TRUE)
```

@@

Search variables:
```{r}
find_census_vectors("income", 
                    dataset = "CA16", 
                    type = "total", 
                    query_type = "keyword", 
                    interactive = FALSE)
```

We need v_CA16_2406 to v_CA16_2425: these are the number of households at certain levels of total income. Retrieve data:
```{r}
data_da_2016 <- get_census(dataset='CA16', 
                          regions=list(CMA=c("35537")),
                          vectors=c("v_CA16_2406",
                                    "v_CA16_2407",
                                    "v_CA16_2408",
                                    "v_CA16_2409",
                                    "v_CA16_2410",
                                    "v_CA16_2411",
                                    "v_CA16_2412",
                                    "v_CA16_2413",
                                    "v_CA16_2414",
                                    "v_CA16_2415",
                                    "v_CA16_2416",
                                    "v_CA16_2417",
                                    "v_CA16_2418",
                                    "v_CA16_2419",
                                    "v_CA16_2420",
                                    "v_CA16_2421",
                                    "v_CA16_2422",
                                    "v_CA16_2423",
                                    "v_CA16_2424",
                                    "v_CA16_2425",
                                    "v_CA16_2426"),
                          level='DA', 
                          use_cache = FALSE,
                          geo_format = "sf")
```

@@

data_da_2016 %>%
  mutate(income_less40k = (`v_CA16_2406: Under $5,000` + 
                             `v_CA16_2407: $5,000 to $9,999` + 
                             `v_CA16_2408: $10,000 to $14,999` +
                             `v_CA16_2409: $15,000 to $19,999` +
                             `v_CA16_2410: $20,000 to $24,999` +  
                             `v_CA16_2411: $25,000 to $29,999` +
                             `v_CA16_2412: $30,000 to $34,999` +
                             `v_CA16_2413: $35,000 to $39,999`)) %>%
  tm_shape() +
  tm_polygons("income_less40k",
              alpha = 0.8)
```

@@

THE NEXT NEEDS TO BE CHANGED TO DO THE ANALYSIS BY INCOME GROUP, NOT BY AGE GROUP

>>>>

We can get an API key to use `cancensus` from [CensusMapper](https://censusmapper.ca/). Once you have an API key, it can be stored locally this way:
```{r eval=FALSE}
set_api_key("YOUR API KEY HERE")
```

@@

Search variables:
```{r}
find_census_vectors("age", 
                    dataset = "CA16", 
                    type = "total", 
                    query_type = "keyword", 
                    interactive = FALSE)
```

We need v_CA16_2406 to v_CA16_2425: these are the number of households at certain levels of total income. Retrieve data:
```{r}
data_da_2016 <- get_census(dataset='CA16', 
                          regions=list(CMA=c("35537")),
                          vectors=c("v_CA16_2406",
                                    "v_CA16_2407",
                                    "v_CA16_2408",
                                    "v_CA16_2409",
                                    "v_CA16_2410",
                                    "v_CA16_2411",
                                    "v_CA16_2412",
                                    "v_CA16_2413",
                                    "v_CA16_2414",
                                    "v_CA16_2415",
                                    "v_CA16_2416",
                                    "v_CA16_2417",
                                    "v_CA16_2418",
                                    "v_CA16_2419",
                                    "v_CA16_2420",
                                    "v_CA16_2421",
                                    "v_CA16_2422",
                                    "v_CA16_2423",
                                    "v_CA16_2424",
                                    "v_CA16_2425",
                                    "v_CA16_2426",
                                    "v_CA16_2397", # median hhinc
                                    # age variables
                                    "v_CA16_10", #<1
                                    "v_CA16_13", #1
                                    "v_CA16_16", #2
                                    "v_CA16_7", #0-4 yo
                                    "v_CA16_25", #5-9 yo
                                    "v_CA16_43", #10-14 yo
                                    "v_CA16_64", # 15-19 yo
                                    "v_CA16_244"), # 65+
                          level='DA', 
                          use_cache = FALSE,
                          geo_format = "sf")
```

#### Prepare Geography

Clip `data_da_2016`:
```{r}
data_da_2016 <- data_da_2016 %>%
  st_intersection(hamilton_cma)
```

#### Prepare Summary Measures

Calculate number of households with incomes < \$40k per DA and proportion of households with incomes < \$40k per DA:
```{r}
data_da_2016 <- data_da_2016 %>%
  filter(`Region Name` == "Hamilton") %>%
  mutate(income_less40k = (`v_CA16_2406: Under $5,000` + 
                             `v_CA16_2407: $5,000 to $9,999` + 
                             `v_CA16_2408: $10,000 to $14,999` +
                             `v_CA16_2409: $15,000 to $19,999` +
                             `v_CA16_2410: $20,000 to $24,999` +  
                             `v_CA16_2411: $25,000 to $29,999` +
                             `v_CA16_2412: $30,000 to $34,999` +
                             `v_CA16_2413: $35,000 to $39,999`),
         p_inc_less40k = income_less40k/Households)
```

@@

data_da_2016 %>%
  tm_shape() +
  tm_polygons("income_less40k",
              alpha = 0.8)
```

Label DAs with income quintiles:
```{r}
data_da_2016 <- data_da_2016 %>%
  rename(Median_household_income = `v_CA16_2397: Median total income of households in 2015 ($)`) %>%
  mutate(income_quintile = case_when(Median_household_income < mhi_q[2] ~ "Bottom 20%",
                                    Median_household_income >= mhi_q[2] & Median_household_income < mhi_q[3] ~ "Fourth 20%",
                                    Median_household_income >= mhi_q[3] & Median_household_income < mhi_q[4] ~ "Third 20%",
                                    Median_household_income >= mhi_q[4] & Median_household_income < mhi_q[5] ~ "Second 20%",
                                    Median_household_income >= mhi_q[5] ~ "Top 20%"),
         income_quintile = factor(income_quintile,
                                levels = c("Top 20%",
                                           "Second 20%",
                                           "Third 20%",
                                           "Fourth 20%",
                                           "Bottom 20%"),
                                ordered = TRUE))
```

Calculate low-income age totals
```{r}
data_da_2016 <- data_da_2016 %>%
  mutate(tot_age_0to2 = `v_CA16_10: Under 1 year` + `v_CA16_13: 1` + `v_CA16_16: 2`,
         linc_age_0to2 = tot_age_0to2 * p_inc_less40k,
         tot_age_0to19 = `v_CA16_7: 0 to 4 years` + `v_CA16_25: 5 to 9 years` + `v_CA16_43: 10 to 14 years` + `v_CA16_64: 15 to 19 years`,
         linc_age_0to19 = tot_age_0to19 * p_inc_less40k,
         tot_age_65plus = `v_CA16_244: 65 years and over`,
         linc_age_65plus = tot_age_65plus * p_inc_less40k,
         linc_population = Population * p_inc_less40k)
```

Plot result
```{r}
tmap_mode(mode = "view")

data_da_2016 %>%
  tm_shape() +
  tm_polygons("linc_population",
              alpha = 0.8)
```

@@
